\bf Measuring background and systemic risk using financial time series

### 2015-06-05 srisk7.tex
GL changes
GL changes in srisk7

\if0\blind
{
  \title{\bf Measuring background and systemic risk using financial time series}
  \author{Piet de Jong\thanks{
    The authors gratefully acknowledge CIFR, the Centre of International Financial Regulation for financial support and other assistance.}\hspace{.2cm}\\
    Department of Applied Finance and Actuarial Studies, Macquarie University\\
    and \\
    Geoff Loudon\\
    Department of Applied Finance and Actuarial Studies, Macquarie University\\
    and\\
    Weihao Choo\\
    Department of Applied Finance and Actuarial Studies, Macquarie University}
  \maketitle
} \fi

@@

Monitoring stresses and systemic risk in the financial system is a key function of macro--prudential regulation. Central banks and other financial regulatory bodies such as the US Financial Stability Oversight Council, the European Systemic Risk Board and the Australian Prudential Regulation Authority [APRA] are concerned with developing and applying early warning quantitative  measures  to warn of potential episodes of heightened systemic risk. Systemic risk has many dimensions. For example, in their influential review paper, \cite{Bisias2012} present a taxonomy of six categories of systemic risk comprising 31 quantitative risk measures. While attempts to gain consensus on a theoretical definition have proved elusive, several empirical measures of systemic risk have been developed and applied with  success.  An important example is the SRISK measure due to \cite{brownlees2015} -- an index that measures the expected capital shortfall of a financial institution conditional on a large and lengthy decline in equity markets. \cite{brownlees2015} show the practical usefulness of their measure in three ways: (i) SRISK rankings identify systemically risky US banks during the global financial crisis [GFC]; (ii) pre-crisis SRISK helps predict capital injections by the Fed Reserve; (iii) aggregate SRISK provides early warning of declines in industrial production and higher unemployment.

Our paper refines and extends the model of \cite{brownlees2015} in several interesting and important ways. First, we reinterpret their SRISK measure as a put on the expected shortfall using stressed expectation. While they condition on the event of a severe market downturn, we take the stressed expectation of the put value. This eNABles us to capture the impact of increased return volatility due to stress. Second, we decompose their SRISK measure into background stress and imposed system stress. Background stress is a function of existing volatility and leverage. As these factors increase, background stress rises and this makes the firm more susceptible to systemic risk, regardless of the stress event itself. Third,    
\cite{brownlees2015} condition on an absolute cutoff of arbitrary size and length. We suggest that it is also helpful to specify conditional behaviour in percentile terms relative to current market conditions. This provides a sharper focus on the incremental impact of further stress and allows us to draw stronger implications of SRISK numbers. These extensions provide additional insights into the sources and economic nature of systemic risk. They provide potentially new and improved ways of anticipating episodes of heightened systemic risk. Their practical use can be expected to help regulators monitor systemic risk in a timely fashion to facilitate remedial action where necessary.

@@

While our measures are designed for use in all financial systems, it is of particular interest to test them using the Australian setting for at least two reasons. First, the Australian financial system is dominated by only four banks. In Australia there are four major banks that hold 78\% of the total assets of all Authorised Deposit-taking Institutions in  Australia.\footnote{Based on balances at 31 December 2014 disclosed in APRA's Quarterly Authorised Deposit-taking Institution [ADI] Performance Report for March 2015.} It is therefore obvious that the four majors are systemically important and the remaining ADIs are not so unless stress in one of these minors somehow creates financial contagion to the majors, in an unforeseeable way. Second, the Australian banking system proved to be resilient through the recent GFC, implying that systemic risk did not reach the extreme levels experienced by other countries where bank failures occurred. Both these features of the Australian system suggest that it may be difficult to find useful additional information about systemic risk by refining quantitative measures like SRISK. Since we are able to show that our measures are informative about systemic risk in Australia, we conjecture that they will provide additional insights elsewhere.

To illustrate the nature and usefulness of our measures, we use publicly available daily financial data for the eight Australian banks detailed in \tref{eightbanks} spanning the period from  3 April 2000 through to 1 December 2014.  The data is described in detail in \aref{data}.  Prices, adjusted for dividends are plotted in \fref{prices}.  Prices are combined with number of shares to derive total equity. Total debt for each firm is also collected from annual accounting reports. % (Should these be plotted? No, not informative;  leverage ratio plots are sufficient).

@@

\section{Literature review}\label{litrev}
Many papers in the finance and economics literature are devoted to the development and application of a variety of quantitative measures of systemic risk or financial stress. \cite{Bisias2012} provide an instructive and qualitative survey of numerous measures in use. \cite{Giglio2015} supply empirical evidence on the ability of many of these measures, both individually and collectively, to provide early warning signals of deterioration in macro-economic conditions. In this section, we provide a selective overview of the measures most closely related to SRISK and its relevance for assessing systemic risk in the financial sector.

The genesis of SRISK lies in a series of papers including \cite{acharya2012capital}, \cite{acharya2012measuring} and \cite{brownlees2010volatility}. In these papers, SRISK is defined as the expected capital shortfall of a financial firm, conditional on a crisis. A crisis is deemed to occur whenever a relevant market index suffers a significant decline over a chosen horizon. \cite{Benoit2013} show how several popular systemic risk measures including MES and SRISK are transformations of market risk measures and derive conditions under which these measures provide similar rankings of systemically important financial institutions.

\cite{brownlees2015} develop SRISK. [***** to add a fuller description of this paper given it is the key paper for our work]

\cite{Engle2015} extend the model in \cite{brownlees2015} to incorporate global, country and industry effects. This extended model makes effective use of \cite{Engle2014dcb} dynamic conditional model. Their empirical methods are able to provide information of the relative systemic importance of industry type and country identity.

\cite{adrian2011covar} propose an alternative measure of systemic risk that emphasises the co-dependence of financial firms and the importance of risk spillovers within the financial sector. They introduce CoVaR$_i$ which computes the Value-at-Risk [VaR] of the entire financial system, conditional on institution $i$ being in distress. Distress is defined as the state where the institution is exactly at its VaR level. \cite{Girardi2013} improve CoVaR by conditioning on the institution being at most at its VaR. This generalisation is useful as it takes into account severity of tail losses and facilitates back-testing of CoVaR.

Empirical estimates of SRISK measure require a model for the joint distribution of returns of individual firms and the market. Many of these papers use \cite{engle2002dynamic}'s GARCH-DCC model aiming to capture the time-varying dynamics of the return (co)variances in a feasible manner. While we also use this model in our empirical analysis, we emphasize that our approach can utilise any appropriate model for simulating the joint distribution of forward firm and market returns.

@@

The above two situations can be combined.   Suppose  $\phi(u)=c(\alpha-u)^{n-1}$ for $u\le \alpha$ and 0 otherwise and where $c$ is such that $\phi(u)$ integrates to 1. Then
$$
\mu_{it}+s_{it} \approx \frac{1}{N/c}\sum_{u^\o_{mt}<\alpha}  \left(\alpha-u^\o_{mt}\right)^{n-1}p_{it}^\o\ .
$$
Similar to the first example, the stressed sample picks up the bottom $\alpha$ of market returns. The bottom $\alpha$ of market returns is further stressed by  progressively overweighing returns as the percentile approaches 0.   Large simulation effort $N$ is required for a reasoNABle approximation since $N/c$ is the effective simulation size and $c$ is small.

>>>>

\if0\blind
{
  \title{\bf Measuring background and systemic risk using financial time series}
  \author{Piet de Jong\thanks{
    The authors gratefully acknowledge CIFR, the Centre of International Financial Regulation for financial support and other assistance.}\hspace{.2cm}\\
    Department of Applied Finance and Actuarial Studies, Macquarie University\\
    and \\
    Geoffrey Loudon\\
    Department of Applied Finance and Actuarial Studies, Macquarie University\\
    and\\
    Weihao Choo\\
    Department of Applied Finance and Actuarial Studies, Macquarie University}
  \maketitle
} \fi

@@

Monitoring stresses and systemic risk in the financial system is a key function of macro--prudential regulation. Central banks and other financial regulatory bodies such as the US Financial Stability Oversight Council, the European Systemic Risk Board and the Australian Prudential Regulation Authority [APRA] are concerned with developing and applying early warning quantitative  measures  to warn of potential episodes of heightened systemic risk. Systemic risk has many dimensions. For example, in their influential review paper, \cite{Bisias2012} present a taxonomy of six categories of systemic risk comprising 31 quantitative risk measures. While attempts to gain consensus on a theoretical definition have proved elusive, several empirical measures of systemic risk have been developed and applied with  success.  An important example is the SRISK measure due to \cite{brownlees2015} -- an index that measures the expected capital shortfall of a financial institution conditional on a large and lengthy decline in equity markets. \cite{brownlees2015} show that their SRISK based measures are able to identify systemically important financial firms and are informative about the impact of financial crises on future macro-economic conditions.

Our paper refines and extends the model of \cite{brownlees2015} in several interesting and important ways. First, we reinterpret their SRISK measure as a put on the expected shortfall using stressed expectation. While they condition on the event of a severe market downturn, we take the stressed expectation of the put value. This enables us to capture the impact of increased return volatility due to stress. Second, we decompose their SRISK measure into background stress and imposed system stress. Background stress is a function of existing volatility and leverage. As these factors increase, background stress rises and this makes the firm more susceptible to systemic risk, regardless of the stress event itself. Third,    
\cite{brownlees2015} condition on an absolute cutoff of arbitrary size and length. We suggest that it is also helpful to specify conditional behaviour in percentile terms relative to current market conditions. This provides a sharper focus on the incremental impact of further stress and allows us to draw stronger implications from SRISK numbers. These extensions provide additional insights into the sources and economic nature of systemic risk. They provide potentially new and improved ways of anticipating episodes of heightened systemic risk. Their practical use can be expected to help regulators monitor systemic risk in a timely fashion to facilitate remedial action where necessary.

@@

While our measures are designed for use in all financial systems, it is of particular interest to test them using the Australian setting for at least two reasons. First, the Australian financial system is dominated by only four banks. In Australia there are four major banks that hold 78\% of the total assets of all Authorised Deposit-taking Institutions in  Australia.\footnote{Based on balances at 31 December 2014 disclosed in APRA's Quarterly Authorised Deposit-taking Institution [ADI] Performance Report for March 2015.} It is therefore obvious that the four majors are systemically important and the remaining ADIs are not so unless stress in one of these minors somehow creates financial contagion to the majors, in an unforeseeable way. Second, the Australian banking system proved to be resilient through the recent global financial crisis [GFC], implying that systemic risk did not reach the extreme levels experienced by other countries where bank failures occurred. Both these features of the Australian system suggest that it may be difficult to find useful additional information about systemic risk by refining quantitative measures like SRISK. Since we are able to show that our measures are informative about systemic risk in Australia, we conjecture that they will provide additional insights elsewhere.

To demonstrate the nature and usefulness of our measures, we use publicly available daily financial data for the eight Australian banks detailed in \tref{eightbanks} spanning the period from  3 April 2000 through to 1 December 2014.  The data is described in detail in \aref{data}.  Prices, adjusted for dividends are plotted in \fref{prices}.  Prices are combined with number of shares to derive total equity. Total debt for each firm is also collected from annual accounting reports. % (Should these be plotted? No, not informative;  leverage ratio plots are sufficient).

@@

\section{Literature review}\label{litrev}
Many papers in the finance and economics literature are devoted to the development and application of a variety of quantitative measures of systemic risk or financial stress. \cite{Bisias2012} provide an instructive and qualitative survey of numerous measures in use. \cite{Giglio2015} supply empirical evidence on the ability of many of these measures, both individually and collectively, to provide early warning signals of deterioration in macro-economic conditions. In this section, we provide a selective overview of those measures most closely related to SRISK and its relevance for assessing systemic risk in the financial sector.

The genesis of SRISK lies in a series of papers including \cite{acharya2012aer}, \cite{acharya2012wp} and \cite{brownlees2010volatility}. In these papers, SRISK is defined as the expected capital shortfall of a financial firm, conditional on a crisis. A crisis is deemed to occur whenever a relevant market index suffers a significant decline over a chosen horizon. SRISK measures require an estimate of the long run marginal expected shortfall [LRMES] which is typically obtained by simulation methods.

\cite{brownlees2015} further develop the empirical methodology to construct SRISK measures. SRISK depends on the firm's size, leverage and its LRMES. The LRMES is obtained by assuming a dynamic process for the joint distribution of firm and market returns. \cite{brownlees2015} use  the standard GARCH-DCC model of \cite{engle2002dynamic} with threshold ARCH volatilities on the basis that this represents a good trade-off between model complexity and prediction accuracy. The LRMES is computed as the Monte Carlo average of simulated multi-period returns,  conditional on the return being worse than the cut-off level chosen to identify a financial crisis. Using a sample of large US financial firms,  \cite{brownlees2015} show the practical usefulness of their measure in three ways: (i) SRISK rankings identify systemically risky US banks during the GFC; (ii) pre-crisis SRISK helps predict capital injections by the Fed Reserve; (iii) aggregate SRISK provides early warning of declines in industrial production and higher unemployment.

\cite{Engle2015} extend the model in \cite{brownlees2015} to incorporate global, country and industry effects. This extended model makes effective use of \cite{Engle2014dcb}'s dynamic conditional beta model. Their empirical methods are able to provide information about the relative systemic importance of industry type and country identity. For instance, they find  that systemic risk in Europe during the GFC was predominantly due to banks while France and the UK recorded the largest country-level SRISK numbers.

\cite{adrian2011covar} propose an alternative measure of systemic risk that emphasises the co-dependence of financial firms and the importance of risk spillovers within the financial sector. They introduce CoVaR$_i$ which computes the Value-at-Risk [VaR] of the entire financial system, conditional on institution $i$ being in distress. Distress is defined as the state where the institution is exactly at its VaR level. \cite{Girardi2013} improve CoVaR by conditioning on the institution being at most at its VaR. This generalisation is useful as it takes into account the severity of tail losses and facilitates back-testing of CoVaR.

\cite{acharya2012aer} relate SRISK and CoVaR and demonstrate that assuming the joint distribution of returns is conditionally normal, SRISK is a more complete measure. \cite{Benoit2013} show how several popular systemic risk measures including MES, CoVaR and SRISK are different transformations of market risk measures and derive those conditions under which they provide similar rankings of systemically important financial institutions.

Empirical estimates of SRISK measure require a model for the joint distribution of returns of individual firms and the market. Many of these papers use \cite{engle2002dynamic}'s GARCH-DCC model aiming to capture the time-varying dynamics of the return (co)variances in a feasible manner. While we also use this model in our empirical analysis, we emphasize that our approach can utilise any appropriate model for simulating the joint distribution of firm and market returns going forward.

@@

The above two situations can be combined.   Suppose  $\phi(u)=c(\alpha-u)^{n-1}$ for $u\le \alpha$ and 0 otherwise and where $c$ is such that $\phi(u)$ integrates to 1. Then
$$
\mu_{it}+s_{it} \approx \frac{1}{N/c}\sum_{u^\o_{mt}<\alpha}  \left(\alpha-u^\o_{mt}\right)^{n-1}p_{it}^\o\ .
$$
Similar to the first example, the stressed sample picks up the bottom $\alpha$ of market returns. The bottom $\alpha$ of market returns is further stressed by  progressively overweighing returns as the percentile approaches 0.   Large simulation effort $N$ is required for a reasonable approximation since $N/c$ is the effective simulation size and $c$ is small.
